Geometric variational problems

### 2020-11-04 NSF2020/Description/specificaim2_ver2.tex
working on numerics for proposal

% ----------------------------------------------------------------------
\subsubsection{Nearly touching bodies}
The trapezoid rule is the workhorse for BIEs since it achieves spectral
accuracy for smooth periodic functions~\cite{tre-wei2014}. However,
nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and such near-contact results in  nearly-singular integrands.
Quadrature methods to address such integrands has received a lot of
attention in two and three dimensions~\cite{alpert, kapur, sidi, duffy,
bruno1, bruno2, davis_1984, graglia_2008, hackbusch_sauter_1994,
jarvenpaa_2003, khayat_2005, schwab_1992, ying_2006, beale1, beale2,
goodman_1990, haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2,
ggq3, helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. For two-dimensional problems, an
elegant quadrature method, which we will call a {\em barycentric
quadrature rule}, is a mildly intrusive modification of the trapezoid
rule relies on analytic function theory~\cite{ioa-pap-per1991}. Since
the Stokes equations can be cast in terms of analytic functions, the
barycentric-quadrature rule has recently been
applied~\cite{bar-wu-vee2015, chi-moo-qua2020}. The PIs plan to extend
this quadrature rule so that it can be applied to the screened Laplace
equation.

Another outstanding numerical issue pertaining to simulating the
self-assembly of amphiphilic particles (such as lipid macromolecles) in
a viscous solvent is the collision between amphiphilic particles. The
hydrophobic attraction potential drives the amphiphilic particles to
move towards each other so to minimize exposure to the solvent. Under
such attraction force the fluid is being squeezed out as particles move
toward each other, and the amphiphilic particles can come into physical
contact in finite time. Such particle collisions in a dense rigid body
suspension is a great challenge and can be a bottleneck in large-scale
simulations. In our previous work a Leonard-Jones potential is used to
keep the particles from overlapping~\cite{Fu2018_SIAM}. Such steep
steric interaction at short ranges introduces great numerical stiffness
that limit the time step size.

Yan {\it et al.} proposed a collision-resolution algorithm within the
framework of boundary integral methods to resolve the non-smooth
many-body dynamics due to collisions by formulating the particle
collision process as a linear complementarity problem with geometric
`non-overlapping' constraint~\cite{Yan2019}. We propose to treat the
particle collisions based on Yan {\it et al.}'s work after the first two
goals (see above) are achieved. A potential complication in combining
the algorithms for particle collision into our QBX integral formulation
is the computational cost for the QBX to refine meshes on particles in
collision.

Once the collision-resolution algorithm is implemented, we will check
the computing performance first before we embark on incorporating
fluctuating hydrodynamics~\cite{Bao17,Bao18}, which might be important
even at the scales of the amphiphilic particle size in some cases. We
propose to extend the high-order time-integration scheme for a
stochastic differential equation~\cite{fu2015pre} to this system.

@@

% ----------------------------------------------------------------------
\subsubsection{A hybrid method for close-to-touching Janus particles}
The HAP model for membranes will inevitably lead to the case where some
Janus particles will be very close to each other. Even though QBX can
handle such case accurately, it may lead to excessively large number of
unknowns in order to resolve the unknown densities at those
close-to-touching points. In~\cite{gan2016sisc}, a hybrid scheme is
developed to treat the close-to-touching spheres for the interface
problem of the Laplace equation. The key idea is to combine the analytic
image method and spectrally accurate method of moments, where the
analytic image method will capture the most singular part of the unknown
density at the close-to-touching points. With the aid of the FMM, the
scheme achieves linear complexity with much less number of unknowns on
the material interfaces. We plan to study the hybrid scheme to treat
the close-to-touching case for both the screened Laplace equation and
the mobility problem to further improve the efficiency of the numerical
algorithm for the HAP model.

>>>>

% ----------------------------------------------------------------------
\subsubsection{Nearly touching bodies}
Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and such near-contact results in nearly-singular integrands
in both two and three dimensions. Quadrature methods to address such
integrands has received a lot of attention in two and three
dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2, davis_1984,
graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003, khayat_2005,
schwab_1992, ying_2006, beale1, beale2, goodman_1990, haroldson_1998,
lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3, helsing_2008a,
helsing_integral_2009, helsing_tutorial_2012, klockner2013jcp, qbx2,
wala2019jcp, af2018sisc, siegel2018jcp, rachh2017jcp, ding2019arxiv,
bar2014}. For two-dimensional problems, the trapezoid rule is the
workhorse for BIEs since it achieves spectral accuracy when integrands
are not nearly-singular~\cite{tre-wei2014}. To address nearly-singular
integrands of two dimensional BIEs, we will use a {\em barycentric
quadrature rule} that only requires a slight modification of the
trapezoid rule~\cite{ioa-pap-per1991}. In its current form, this method
requires the layer potential satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}. The PIs will extend
this quadrature rule so that it can be applied to the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log\|\mathbf{x} - \mathbf{y}\|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log\|\mathbf{x} - \mathbf{y}\|$. Using this
decomposition, the layer potential involving $K_1(\mathbf{x}\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and the layer potential
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all
$\mathbf{x}$ and $\mathbf{y}$.

Even with accurate quadrature and time stepping methods (see
Section~\ref{subsec:NumericalIssues}) methods, rigid body simulations
can still result in unphysical contact. In our previous work a
Leonard-Jones potential is used to keep the particles from
overlapping~\cite{Fu2018_SIAM}. Such steep steric interaction at short
ranges introduces great numerical stiffness that limit the time step
size.

\textcolor{red}{
Another outstanding numerical issue pertaining to simulating the
self-assembly of amphiphilic particles (such as lipid macromolecles) in
a viscous solvent is the collision between amphiphilic particles. The
hydrophobic attraction potential drives the amphiphilic particles to
move towards each other so to minimize exposure to the solvent. Under
such attraction force the fluid is being squeezed out as particles move
toward each other, and the amphiphilic particles can come into physical
contact in finite time. Such particle collisions in a dense rigid body
suspension is a great challenge and can be a bottleneck in large-scale
simulations. In our previous work a Leonard-Jones potential is used to
keep the particles from overlapping~\cite{Fu2018_SIAM}. Such steep
steric interaction at short ranges introduces great numerical stiffness
that limit the time step size. 
}

\textcolor{red}{
Yan {\it et al.} proposed a collision-resolution algorithm within the
framework of boundary integral methods to resolve the non-smooth
many-body dynamics due to collisions by formulating the particle
collision process as a linear complementarity problem with geometric
`non-overlapping' constraint~\cite{Yan2019}. We propose to treat the
particle collisions based on Yan {\it et al.}'s work after the first two
goals (see above) are achieved. A potential complication in combining
the algorithms for particle collision into our QBX integral formulation
is the computational cost for the QBX to refine meshes on particles in
collision.
}

\textcolor{red}{
Once the collision-resolution algorithm is implemented, we will check
the computing performance first before we embark on incorporating
fluctuating hydrodynamics~\cite{Bao17,Bao18}, which might be important
even at the scales of the amphiphilic particle size in some cases. We
propose to extend the high-order time-integration scheme for a
stochastic differential equation~\cite{fu2015pre} to this system.
}

@@

% ----------------------------------------------------------------------
\subsubsection{A hybrid method for close-to-touching Janus particles}
\textcolor{red}{
The HAP model for membranes will inevitably lead to the case where some
Janus particles will be very close to each other. Even though QBX can
handle such case accurately, it may lead to excessively large number of
unknowns in order to resolve the unknown densities at those
close-to-touching points. In~\cite{gan2016sisc}, a hybrid scheme is
developed to treat the close-to-touching spheres for the interface
problem of the Laplace equation. The key idea is to combine the analytic
image method and spectrally accurate method of moments, where the
analytic image method will capture the most singular part of the unknown
density at the close-to-touching points. With the aid of the FMM, the
scheme achieves linear complexity with much less number of unknowns on
the material interfaces. We plan to study the hybrid scheme to treat
the close-to-touching case for both the screened Laplace equation and
the mobility problem to further improve the efficiency of the numerical
algorithm for the HAP model.
}
